---
output: github_document
bibliography: README.bib
csl: CJFAS.csl
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbbt

The goal of **rbbt** is to connect R to the [Better Bibtex for Zotero connector](https://retorque.re/zotero-better-bibtex/). This allows the insertion of in-text citations (pandoc or LaTex style) and BibLaTex bibliography items directly into the RStudio editor using the RStudio addin, or to the console otherwise.

## Installation

You can install rbbt from GitHub with:

``` r
remotes::install_github("paleolimbot/rbbt")
```

You will also need [Zotero](https://www.zotero.org/) installed and running, and the [Better BibTeX for Zotero](https://retorque.re/zotero-better-bibtex/installation/) add-on installed.

## Example

Insert bibliography text using your Better BibTex citation keys:

```{r, comment=""}
# uses the citation keys you've defined with BetterBibTex
rbbt::bbt_bib("dunnington_etal18")
```

Insert bibliography text from selected items in Zotero using the RStudio Addin, or print to the console using `bbt_bib_selected()`:

```{r, comment=""}
# uses whatever is currently selected in the zotero window
rbbt::bbt_bib_selected()
```

To make this work seamlessly with [knitr]() and [rmarkdown](), you can use `bbt_write_bib()` to detect citations in the current document and write your bibliography file, all in one go! One way to go about this is to use something like this in your YAML front matter:

```{r, echo=FALSE, comment=""}
cat("
---
title: \"Zotero + Better BibTeX + RMarkdown using rbbt\"
output: html_document
bibliography: \"`r rbbt::bbt_write_bib('biblio.json', overwrite = TRUE)`\"
---
")
```

This is still experimental, so file an issue if it fails!
