---
output: github_document
bibliography: README.bib
csl: CJFAS.csl
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbbt

<!-- badges: start -->
[![R build status](https://github.com/paleolimbot/rbbt/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/rbbt/actions)
<!-- badges: end -->

The goal of rbbt is to connect R to the [Better Bibtex for Zotero connector](https://retorque.re/zotero-better-bibtex/). This allows the insertion of in-text citations (pandoc or LaTex style) and BibLaTex bibliography items directly into the RStudio editor using the RStudio addin, or to the console otherwise.

## Installation

You can install rbbt from GitHub with:

``` r
remotes::install_github("paleolimbot/rbbt")
```

```{r}
library(rbbt)
```

You will also need [Zotero](https://www.zotero.org/) installed and running, and the [Better BibTeX for Zotero](https://retorque.re/zotero-better-bibtex/installation/) add-on installed.

## RStudio Addin

This package is most useful for the RStudio addins that insert citations and references into the editor. The "Insert Zotero Citation" will pop up a zotero search where you can search for a reference to add to your writing (Markdown, RMarkdown, or LaTeX). The "Insert Zotero Bibliography" addin inserts the bibliographical information corresponding to the selected item in the Zotero window. Finally, "Update bibliography for current document" runs `bbt_update_bib()` to update the bibliography based on the .Rmd file currently in the editor window. You can bind either of these to keyboard shortcuts in RStudio. The addins may behave differently depending on which file you have open (e.g., if you have a .bib file open, it will use the biblatex translator)...if you need more fine-grained control, try one of the console functions (like `bbt_bib()`!).

## RStudio shortcuts

In RStudio, it is possible to assign a keyboard shortcut for **rbbt** addins. Just go to *Tools --> Modify Keyboard Shortcuts* and a small window will open with all the available commands to assign a keyboard shortcut. Then, using the **Scope** column, find the *Addin* type and configure the shortcuts for **rbbt** commands. For example: to set a shortcut for `bbt_update_bib()`, go to the row assigned for **"Update bibliography for current document from Zotero "**, click on the **Shortcut** column and press the desired key combination. It is recommended to define shortcuts for **"Update bibliography for current document from Zotero "** and **"Insert Zotero citation "** as they will be the most used.

It should be noted that RStudio comes configured with some keyboard shortcuts, so it should be avoided that the newly defined shortcuts overlap (if this happens, a warning icon will be displayed in the shortcut assignment window).

Also, for the commands to work correctly, they must be executed having the Rmd or Qmd file in the active tab in RStudio. That is to say, both addins mentioned above will detect the citations and the bibliography file from the active tab (from the main .Rmd or .Qmd file) from where the shortcut is executed, so you must make sure to have that document active when executing the commands, otherwise you will receive the following error message: **"Currently selected editor is not a .qmd, .rmd or .Rmd file "**.

In RStudio, it is possible to assign a keyboard shortcut for rbbt addins. Just go to Tools --> Modify Keyboard Shortcuts and a small window will open with all the available commands to assign a keyboard shortcut. Then, using the **Scope** column, find the **Addin** type and configure the shortcuts for **rbbt** commands. For example: to set a shortcut for `bbt_update_bib()`, go to the row assigned for **"Update bibliography for current document from Zotero "**, click on the **Shortcut** column and press the desired key combination. It is recommended to define shortcuts for **"Update bibliography for current document from Zotero "** and **"Insert Zotero citation "** as they will be the most used.

## knitr + rmarkdown integration

To make this work seamlessly with [knitr](https://yihui.org/knitr/) and [rmarkdown](https://rmarkdown.rstudio.com/), use the "Update bibliography for current document" addin to write your bibliography file based on the citations in the currently selected document. You can use `bbt_update_bib("some_file.Rmd")` to detect citations and write your bibliography file from the console.

# Versions

## 0.0.1
* Add functionality for citing R packages. Now, if the user wants to cite a R package, could indicate it in the Rmd/Qmd file with the syntax `@rpkg_NameOfPackage` (e.g. `@rpkg_rbbt`). `bbt_detect_citations` and `bbt_update_bib` are nor prepared for identify this syntax and adding to the bibliography using `citation` and `toBibtex` tools.
* R native pipe is used in some functions so now it is necessary to run **rbbt** in **R >= 4.1.0**.
* Subtle corrections and improvements in code and documentation.

## 0.0.0.9002

* Changes in `bbt_detect_citations` in order to avoid including citations that starts whether with *fig-* or *tbl-* (associtaed with Quarto syntaxis).
* Minor improvements in definiton of external functions and dependencies.
