---
output: github_document
bibliography: README.bib
csl: CJFAS.csl
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbbt

<!-- badges: start -->
[![R build status](https://github.com/paleolimbot/rbbt/workflows/R-CMD-check/badge.svg)](https://github.com/paleolimbot/rbbt/actions)
<!-- badges: end -->

The goal of rbbt is to connect R to the [Better Bibtex for Zotero connector](https://retorque.re/zotero-better-bibtex/). This allows the insertion of in-text citations (pandoc or LaTex style) and BibLaTex bibliography items directly into the RStudio editor using the RStudio addin, or to the console otherwise.

## Installation

You can install rbbt from GitHub with:

``` r
remotes::install_github("paleolimbot/rbbt")
```

```{r}
library(rbbt)
```

You will also need [Zotero](https://www.zotero.org/) installed and running, and the [Better BibTeX for Zotero](https://retorque.re/zotero-better-bibtex/installation/) add-on installed.

## RStudio Addin

This package is most useful for the RStudio addins that insert citations and references into the editor. The "Insert Zotero Citation" will pop up a zotero search where you can search for a reference to add to your writing (Markdown, RMarkdown, or LaTeX). The "Insert Zotero Bibliography" addin inserts the bibliographical information corresponding to the selected item in the Zotero window. You can bind either of these to keyboard shortcuts in RStudio. The addin behaves differently depending on which file you have open (e.g., if you have a .bib file open, it will use the biblatex translator)...if you need more fine-grained control, try one of the console functions (like `bbt_bib()`!).

## knitr + rmarkdown integration

To make this work seamlessly with [knitr](https://yihui.org/knitr/) and [rmarkdown](https://rmarkdown.rstudio.com/), you can use `bbt_write_bib()` to detect citations in the current document and write your bibliography file, all in one go! One way to go about this is to use something like this in your YAML front matter:

```{r, echo=FALSE, comment=""}
cat("
---
title: \"Zotero + Better BibTeX + RMarkdown using rbbt\"
output: html_document
bibliography: \"`r rbbt::bbt_write_bib('biblio.json', overwrite = TRUE)`\"
---
")
```

This is still experimental, so file an issue if it fails! You can use `bbt_write_bib("biblio.json", bbt_detect_citations("file.Rmd"))` to do this manually.

## Programmatically fetch bibliography info

Fetch bibliography text using a list of Better BibTex citation keys:

```{r, comment=""}
# uses the citation keys you've defined with Better BibTeX
rbbt::bbt_bib("dunnington_etal18", translator = "biblatex", .action = bbt_print)
```

Fetch bibliography text from selected items in Zotero:

```{r, comment=""}
# uses whatever is currently selected in the zotero window
rbbt::bbt_bib_selected(translator = "biblatex", .action = bbt_print)
```
